---
title: "Serology Analysis Survey"
output: html_notebook
---

Load up the packages to perform the analysis

```{r}
library(dplyr)
library(tibble)
library(tidyr)
library(survey)
library(ggplot2)
library(apyramid)
library(srvyr)
```



Loading the datasets

```{r}
df_population <- readRDS("/conf/EAVE/GPanalysis/data/EAVE_demographics_SK.rds")
df_serology_pc <- readRDS("/conf/EAVE/GPanalysis/data/serology_primcare_march22.rds")
nrow(df_population)
nrow(df_serology_pc)
```

Create population dataframe with factors for different age groups and sex:

```{r}
df_population <- df_population %>%
  ## ensure that age and sex are factors
  mutate(age_group = cut(ageYear, breaks = c(0,10,20,30,40,50,60,70,80,90,1000),
                             label=c("0-9", 
                                       "10-19", 
                                       "20-29",
                                       "30-39", 
                                       "40-49",
                                       "50-59",
                                       "60-69",
                                       "70-79",
                                       "80-89",
                                       "90+"))) %>%
  mutate(sex = factor(Sex))
nrow(df_population)
```

Create groupings for the different populations:

```{r}
source_population <- df_population %>%
                     group_by(age_group, sex) %>% 
                     summarise(population = n()) %>% 
                     drop_na()
source_population
```



```{r}
max_prop <- 11     # choose the highest proportion you want to show 
step <- 5           # choose the space you want beween labels 

## this part defines vector using the above numbers with axis breaks
breaks <- c(
    seq(max_prop/100 * -1, 0 - step/100, step/100), 
    0, 
    seq(0 + step / 100, max_prop/100, step/100)
    )

## this part defines vector using the above numbers with axis limits
limits <- c(max_prop/100 * -1, max_prop/100)

## this part defines vector using the above numbers with axis labels
labels <-  c(
      seq(max_prop, step, -step), 
      0, 
      seq(step, max_prop, step)
    )


p1 <- source_population %>%
  ungroup() %>% 
  mutate(proportion = population / sum(population)) %>% 
  ## plot pyramid 
  age_pyramid(
            age_group = age_group, 
            split_by = sex, 
            count = proportion, 
            proportional = TRUE) +
  ## only show the y axis label (otherwise repeated in all three plots)
  labs(title = "Source population", 
       y = "", 
       x = "Age group (years)") + 
  ## make the x axis the same for all plots 
  scale_y_continuous(breaks = breaks, 
    limits = limits, 
    labels = labels)
p1
```



Make a new dataframe that is the unique population in the serology samples


```{r}
survey_data <- df_serology_pc %>% distinct(EAVE_LINKNO, .keep_all = TRUE)  %>% select(EAVE_LINKNO,age,sex) %>%
                      mutate(sex = factor(sex),
                             age_group = cut(age, breaks = c(0,10,20,30,40,50,60,70,80,90,120),
                             label=c("0-9", 
                                       "10-19", 
                                       "20-29",
                                       "30-39", 
                                       "40-49",
                                       "50-59",
                                       "60-69",
                                       "70-79",
                                       "80-89",
                                       "90+"))) 
nrow(survey_data)
```


```{r}
max_prop <- 12     # choose the highest proportion you want to show 
step <- 5           # choose the space you want beween labels 

## this part defines vector using the above numbers with axis breaks
breaks <- c(
    seq(max_prop/100 * -1, 0 - step/100, step/100), 
    0, 
    seq(0 + step / 100, max_prop/100, step/100)
    )

## this part defines vector using the above numbers with axis limits
limits <- c(max_prop/100 * -1, max_prop/100)

## this part defines vector using the above numbers with axis labels
labels <-  c(
      seq(max_prop, step, -step), 
      0, 
      seq(step, max_prop, step)
    )


p2  <- age_pyramid(survey_data, 
                 age_group = "age_group", 
                 split_by = "sex",
                 proportion = TRUE) + 
  ## only show the x axis label (otherwise repeated in all three plots)
  labs(title = "Unweighted sample population", 
       y = "Proportion (%)", 
       x = "") + 
  ## make the x axis the same for all plots 
  scale_y_continuous(breaks = breaks, 
    limits = limits, 
    labels = labels)
p2
```

Define a function, stolen from https://rdrr.io/github/R4EPI/epikit/src/R/add_weights_strata.R to obtain weights

```{r}
add_weights_strata <- function(x, p, ...,
                               population = population,
                               surv_weight = "surv_weight",
                               surv_weight_ID = "surv_weight_ID") {

  ## define vars and throw error if not existant
  pop <- tidyselect::vars_select(names(p), {{population}}, .strict = FALSE)
  if (length(pop) == 0) {
    cll   <- match.call()
    ppltn <- rlang::as_name(rlang::enquo(population))
    stop(glue::glue("{ppltn} is not one of the columns of {cll[['p']]}, check spelling"))
  } else {
    population      <- rlang::ensym(pop)
  }
  surv_weight_ID  <- rlang::sym(surv_weight_ID)
  surv_weight     <- rlang::sym(surv_weight)
  onames          <- names(x)

  # create a merger ID by age group and sex
  p <- tidyr::unite(p, !!surv_weight_ID, ...)
  x <- tidyr::unite(x, !!surv_weight_ID, ..., remove = FALSE)

  # get study sample counts
  counts <- dplyr::count(x, !!surv_weight_ID)

  # merge counts to population data
  p <- dplyr::left_join(p, counts, by = rlang::as_name(surv_weight_ID))

  # create weight variable
  p <- dplyr::mutate(p, !!surv_weight := !!population / .data$n)
  p <- dplyr::select(p, !!surv_weight, !!surv_weight_ID)


  # merge to study sample
  res <- dplyr::left_join(x, p, by = rlang::as_name(surv_weight_ID))

  # return in original order of x
  dplyr::select(res, onames, !!surv_weight_ID, !!surv_weight)

}

```

Use the function to add weights 

```{r}
survey_data_with_weights <- add_weights_strata(x = survey_data,
                                    p = source_population,
                                    surv_weight = "surv_weight_strata",
                                    surv_weight_ID = "surv_weight_ID_strata",
                                    age_group, sex) %>%
                                    drop_na(surv_weight_strata)
head(survey_data_with_weights %>% select(-EAVE_LINKNO),5)
```



```{r}
survey_design <- survey_data_with_weights %>% as_survey_design(ids = 1,      # cluster ids
                   weights = surv_weight_strata, # weight variable created above
                   strata = NULL 
                  )
```



```{r}
p3 <- survey_design %>% 
  ## make sure the variables are factors
  mutate(age_group = factor(age_group), 
         sex = factor(sex)) %>%
  age_pyramid(
    age_group = "age_group",
    split_by = "sex", 
    proportion = TRUE) +
  ## only show the x axis label (otherwise repeated in all three plots)
  labs(title = "Weighted sample population", 
       y = "", 
       x = "")  + 
  ## make the x axis the same for all plots 
  scale_y_continuous(breaks = breaks, 
    limits = limits, 
    labels = labels)
p3
```


```{r}
require(gridExtra)
grid.arrange(p1, p2, p3, ncol=2,nrow=2)
```



```{r}
base_survey_design_strata <- svydesign(ids = ~1,  # 1 for no cluster ids
                   weights = ~surv_weight_strata, # weight variable created above
                   data = survey_data_with_weights         
                  )
base_survey_design_strata
```



Try out the mean 

```{r}
svymean( survey_data_with_weights ,
         design = base_survey_design_strata)
```

