---
title: "Serology Data Exploration March 2022"
output: html_notebook
---
## Datasets


Setup dependancies
```{r}
library(dplyr)
library(tibble)
library("tidyr")
library(ggplot2)
require(scales) 
```


### Serology Primary Care 

Load serology Primary Care and print the number of rows of data there are. Also, extract the IgG quantative result as a number.
```{r}

df_serology_pc <- readRDS("/conf/EAVE/GPanalysis/data/serology_primcare_march22.rds") %>% 
                  as_tibble() %>%
                  dplyr::mutate(IgG = readr::parse_number(test_result_quant))

nrow(df_serology_pc)
```

See the column names...
```{r}
colnames(df_serology_pc)
```

Make a quick histogram plot of the number of samples collected over time
```{r}
qplot(df_serology_pc$Sampledate_iso, geom="histogram", fill=I("red"), xlab='Sample Date', ylab='Number of Collected Samples', bins=30) 
```



Count how many times each person has had a sample taken, and then count how many occurrences there are of people with `nsamples`
```{r}
df_serology_pc %>% dplyr::group_by(EAVE_LINKNO) %>% dplyr::summarise(nsamples = dplyr::n()) %>% 
                   dplyr::group_by(nsamples) %>% dplyr::summarise(noccurrences = dplyr::n())
```



Filter the dataframe to get the number of rows where there has been a positive test result
```{r}

df_serology_pc_pos <- df_serology_pc %>% dplyr::filter(test_result_qual == "Positive")
nrow(df_serology_pc_pos)
```


### Serology SNBTS (blood donors)


```{r}
df_serology_bd <- readRDS("/conf/EAVE/GPanalysis/data/serology_snbts_march22.rds") %>% 
                  as_tibble() %>%
                  dplyr::mutate(IgG = readr::parse_number(test_result_quant))

nrow(df_serology_bd)

```

Plot the number of samples collected over time for the blood donors
```{r}
qplot(df_serology_bd$Sampledate_iso, geom="histogram", fill=I("red"), xlab='Sample Date', ylab='Number of Collected Samples', bins=40) 
```


Again, filter the dataframe to get the number of rows where there has been a positive test result:
```{r}

df_serology_bd_pos <- df_serology_bd %>% dplyr::filter(test_result_qual == "Positive")
nrow(df_serology_bd_pos)
```

Check the number of repeat measurements of a person.. 
```{r}
df_serology_bd %>% dplyr::group_by(EAVE_LINKNO) %>% dplyr::summarise(nsamples = dplyr::n()) %>% 
                   dplyr::group_by(nsamples) %>% dplyr::summarise(noccurrences = dplyr::n())
```


### QCOVID (Feb22 Update)

Here we load the QCOVID dataframe and filter it on those EAVE studies that are present in the serology datasets.

```{r}
df_qcovid <- readRDS("/conf/EAVE/GPanalysis/data/cleaned_data/QCOVID_feb22.rds") %>% 
             as_tibble() 

df_qcovid_pc <- df_qcovid %>% dplyr::filter(EAVE_LINKNO %in% df_serology_pc$EAVE_LINKNO)
nrow(df_qcovid_pc)

df_qcovid_bd <- df_qcovid %>% dplyr::filter(EAVE_LINKNO %in% df_serology_bd$EAVE_LINKNO)
nrow(df_qcovid_bd)

```


```{r}
df_ana_pc <- df_serology_pc %>% dplyr::left_join(df_nrisks) %>% 
          dplyr::select(c("EAVE_LINKNO","Sampledate_iso","test_result_qual","IgG","n_risk_gps")) %>%
          dplyr::mutate(n_risk_gps = ifelse(is.na(n_risk_gps), "0", levels(n_risk_gps)[n_risk_gps]))
          

df_ana_pc_pos <- df_ana_pc %>% dplyr::filter(test_result_qual == "Positive")
nrow(df_ana_pc_pos)
```

Plot a histogram showing the linking of the serology data (primary care) with QCOVID

```{r}
bw <- 50
n <- nrow(df_ana_pc_pos)
p <- ggplot(df_ana_pc_pos, aes(x=IgG, fill=n_risk_gps)) +
     geom_histogram(binwidth = bw) + xlim(0,1000+bw) +
     labs(x="Primary Care IgG", y="Number of Samples", fill="Number of Risks (QCOVID)") +
     theme(aspect.ratio = 0.5) + 
     scale_fill_brewer(palette="Set2")
p
```

Plot a histogram showing the linking of the serology data with QCOVID
```{r}
df_ana_bd <- df_serology_bd %>% dplyr::left_join(df_nrisks) %>% 
          dplyr::select(c("EAVE_LINKNO","Sampledate_iso","test_result_qual","IgG","n_risk_gps")) %>%
          dplyr::mutate(n_risk_gps = ifelse(is.na(n_risk_gps), "0", levels(n_risk_gps)[n_risk_gps]))
          

df_ana_bd_pos <- df_ana_bd %>% dplyr::filter(test_result_qual == "Positive")
```
```{r}
min(df_ana_bd_pos$IgG) 
max(df_ana_bd_pos$IgG) 
```


```{r}
bw <- 0.1
n <- nrow(df_ana_bd_pos)
p <- ggplot(df_ana_bd_pos, aes(x=IgG, fill=n_risk_gps)) +
     geom_histogram(binwidth = bw) + xlim(0,11+bw) +
     labs(x="Blood Donors IgG", y="Number of Samples", fill="Number of Risks (QCOVID)") +
     theme(aspect.ratio = 0.5) + 
     scale_fill_brewer(palette="Set2")
p

```