---
title: "Serology - Fail to Mount analysis"
output: html_document
---


Load some libraries...
```{r}
library(tibble)
library(dplyr)
library(ggplot2)
```


Get the serology PC data
```{r}
df_serology_pc <- readRDS("/conf/EAVE/GPanalysis/data/serology_primcare_march22.rds")
df_serology_pc
```

Also get the qcovid, remove the Age/Sex as we well get them from serology (or better, from demographics at a latter point)
```{r}
df_qcovid <- readRDS("/conf/EAVE/GPanalysis/data/cleaned_data/QCOVID_feb22.rds") %>% select(-Age,-Sex) 
df_qcovid
```

Test out the numbers of people with Diabetes type 1 (seems high?)
```{r}
df_qcovid %>% group_by(Q_DIAG_DIABETES_1) %>% summarise(n=n())
```

Convert QCOVID outcomes into numbers, for some reason many are chars
```{r}
df_qcovid <- df_qcovid %>% mutate(across(starts_with("Q_"),.fns =as.numeric))
```

Load the vaccine data (for those found in serology)
```{r}
df_vac <- readRDS("/conf/EAVE/GPanalysis/data/temp/sero_vacc_time.rds")
df_vac
```

Create a dataframe to work with for analysis
* do some mutate to get the variables in a more convienent form / have better names
* filter for people who have had their 2nd vaccine
* create some new variables for the number of days the measurements were taken after vaccines
* further filter, only take measurements >15 days after 2nd vac
* group by EAVE_LINKNO to look for multiple tests
   * arrange so Positive tests appear first in the group
   * select the first test in the group
   * this is a way to make sure we're dealing with people who have always tested negative
* shortern the dataframe, only keeping some data columns
* join with the qcovid dataframe

```{r}
df <- df_serology_pc %>% 
      left_join(df_vac %>% filter(serology_source=='primary_care_serology')) %>%
      mutate(IgG = readr::parse_number(test_result_quant),
                    SpecimenDate=as.Date(SpecimenDate),
                    QualResult=test_result_qual) %>% 
      filter(Sampledate_iso >= d2_date)  %>% 
      mutate(days_since_vac2= as.numeric(SpecimenDate - d2_date,units='days'),
             days_since_vac3=as.numeric(SpecimenDate - d3_date,units='days')) %>% 
      mutate(days_since_vac3=if_else(days_since_vac3>0,days_since_vac3,0))%>%
      filter(days_since_vac2>15) %>%
      group_by(EAVE_LINKNO) %>% arrange(desc(QualResult)) %>% filter(row_number() ==1) %>% ungroup() %>% 
      select(EAVE_LINKNO,SpecimenDate,QualResult,IgG,days_since_vac2,days_since_vac3,sex,age,d2_product) %>% 
      left_join(df_qcovid)

df
```


Clean-up the memory
```{r}
rm(df_qcovid)
rm(df_vac)
rm(df_serology_pc)
```


Check the column names
```{r}
colnames(df)
```


Create a new dataframe that can be used with GLM .. need to get some char variables as numbers..
```{r}
df_ana <- df %>% filter(!is.na(n_risk_gps)) %>%
          mutate(n_risk_gps_int =  as.factor(as.integer(factor(n_risk_gps,labels=c(1,2,3,4,5)))),
                 outcome = ifelse(QualResult=='Negative',1,0),
                 outcome_pos = ifelse(IgG>1500,1,0),
                 outcome_neg = ifelse(IgG<5,1,0),
                 sex = as.factor(ifelse(sex=='M',1,2))) %>% 
         mutate(product = as.integer(factor(d2_product,labels=c(1,2,3))),
                d2_product = factor(d2_product),
                risk0 = ifelse(n_risk_gps_int==1,1,0),
                risk1 = ifelse(n_risk_gps_int==2,1,0),
                risk2 = ifelse(n_risk_gps_int==3,1,0),
                risk3_4 = ifelse(n_risk_gps_int==4,1,0),
                risk5p = ifelse(n_risk_gps_int==5,1,0)) %>%
        mutate(age_group = cut(age,right=FALSE, breaks = c(0,20,40,60,80,1000),
                               label=c("0-20", 
                                       "20-40", 
                                       "40-60",
                                       "60-80", 
                                       "80+"
                                       ))) %>%
         select(-QualResult)
df_ana
```

Get some numbers for the outputs (0 - mount a response, 1- fail to mount)
```{r}
df_ana %>% group_by(outcome) %>% summarise(n=n())
```

Check these outcomes for different vaccine products
```{r}
df_ana %>% group_by(d2_product,outcome) %>%summarise(n=n())
```


```{r}
library(tidyr)
```
Make a little summary to see which QCOVID variables have at least one outcome that is 1
```{r}
temp <- df_ana %>% filter(outcome_pos==1) %>% select(contains("Q_DIA")) %>% summarise(across(where(is.numeric), ~ sum(.x, na.rm = TRUE))) %>% gather() %>% 
     filter(value>3)
temp$key

```

```
n_risk_gps_int
age
```
```{r}
formula <- paste0("outcome_pos ~ days_since_vac2  + age  + sex + product + Q_BMI + ",paste0(temp$key,collapse = " + "))
formula 
```

Peform the GLM
```{r}


glmFit <- glm(formula,family=poisson,data=df_ana)
summary(glmFit)
```

Create a table with odds ratios and confidence intervals
```{r}

lookup <- c(
  "days_since_vac2"="Days Since 2nd Vaccination", 
  "n_risk_gps_int2"="1 Risk",
  "n_risk_gps_int3"="2 Risk",    
  "n_risk_gps_int4"="3-4 Risk",     
  "n_risk_gps_int5"="5+ Risk",    
  "age"="Age",
  "age_group"="Age",
  "age_group20-40"="Age 20-40",
  "age_group40-60"="Age 40-60",
  "age_group60-80"="Age 60-80",
  "age_group80+"="Age 80+",
  "sex2"="Sex (Female)",
  "product"="Vaccine (AstraZeneca)",
  "Q_BMI"="BMI",
  "Q_DIAG_DIABETES_1"="Diabetes Type-I",  
  "Q_DIAG_RESP_CANCER"="Respitory Cancer",
  "Q_DIAG_ASTHMA"="Asthma",    
  "Q_DIAG_BLOOD_CANCER"="Blood Cancer",
  "Q_DIAG_CHD"="Coronary Heart Disease ",
  "Q_DIAG_COPD"="Chronic Obstructive Pulmonary Disease",   
  "Q_DIAG_CKD_LEVEL"="Chronic Kidney Disease",
  "Q_DIAG_AF"="Atrial Fibrillation",
  "Q_DIAG_CCF"="Heart Failure",
  "Q_DIAG_EPILEPSY"="Epilepsy",     
  "Q_DIAG_FRACTURE"="A prior fracture of hip, wrist, spine or humerus",     
  "Q_DIAG_IMMU"="Immune Deficiency",         
  "Q_DIAG_NEURO"="Neurone Disease",       
  "Q_DIAG_PARKINSONS"="Parkinsons",  
  "Q_DIAG_PULM_RARE"="Cystic Fibrosis or Bronchiectasis or Alveolitis",   
  "Q_DIAG_PVD"="Peripheral Vascular Disease",         
  "Q_DIAG_RA_SLE"="Rheumatoid Arthritis",      
  "Q_DIAG_SEV_MENT_ILL"="Severe Mental Health Illness",
  "Q_DIAG_STROKE"="Stroke",      
  "Q_DIAG_VTE"="Thrombosis or Pulmonary Embolus",
  "Q_DIAG_CEREBRALPALSY"="Cerebralpalsy",
"Q_DIAG_CIRRHOSIS"="Cirrhosis",
"Q_DIAG_CONGEN_HD"="Congenital Heart Disease",
"Q_DIAG_HIV_AIDS"="HIV/Aids",
"Q_DIAG_PULM_HYPER"="Pulmonary Hypertension",
"Q_DIAG_DEMENTIA"="Dementia"
)

tab <- exp(cbind(coef(glmFit), confint(glmFit))) %>% as.data.frame
tab <- tab[-1,]
tab$names <- rownames(tab)
names(tab)[1] <- 'OR'
names(tab)[2] <- 'LCL'
names(tab)[3] <- 'UCL'

tab <- tab %>% mutate(key = as.character(lookup[names]))
tab


```


Make a plot
```{r}
p<- ggplot(tab, aes(x=reorder(key,OR), y=OR)) + 
  geom_pointrange(aes(ymin=LCL, ymax=UCL), width=.2,fill='red',shape=21,
                 position=position_dodge(.9)) +
  ylim(0., 16.0) +
  coord_flip() +
  labs(title='',x='',y='OR (95% CI)') +
  geom_hline(yintercept=c(1), linetype="dotted") +
   scale_y_log10() +
  theme_classic()
p
```