### Setup

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(phsstyles)
library(lemon)
library(eavehelpers)
```

### Load Data

Use the helper function to load up the recommended, stable datasets needed for this analysis
```{r}
eave.data <- eavehelpers::load_recommended_datasets() 
labels(eave.data)
```

create two dataframes that join the two serology datasets up with other EAVE-II files
```{r}
data.pc <- eavehelpers::serology_vaccine_analysis.create_dataframe(eave.data,'serology_primary_care')
data.bd <- eavehelpers::serology_vaccine_analysis.create_dataframe(eave.data,'serology_blood_donors')
```

```{r}
attributes(data.pc)$meta
```


```{r}
attributes(data.bd)$meta
```



```{r}
saveRDS(data.pc,"/conf/EAVE/GPanalysis/data/temp/fail_to_mount_pc_3.rds")
saveRDS(data.bd,"/conf/EAVE/GPanalysis/data/temp/fail_to_mount_bd_3.rds")
```


```{r}
data.pc <- readRDS("/conf/EAVE/GPanalysis/data/temp/fail_to_mount_pc_3.rds")
data.bd <- readRDS("/conf/EAVE/GPanalysis/data/temp/fail_to_mount_bd_3.rds")
print (nrow(data.pc))
print (nrow(data.bd))
```


```{r}
output_folder <- file.path("results", Sys.Date())
dir.create(output_folder, showWarnings = TRUE)
```


```{r}
ggplot(data.pc , aes(x=IgG,fill=as.factor(QualResult))) +
  #geom_histogram(position = "stack", bins=30) +
  geom_histogram(position = "stack", bins=30, data=subset(data.pc, QualResult == 'Negative'), aes(fill='Negative')) +
  #geom_histogram(position = "stack", bins=30, data=subset(data.pc, QualResult == 'Equivocal'), aes(fill='Equivocal')) +
  geom_histogram(position = "stack", bins=30, data=subset(data.pc, QualResult == 'Positive'), aes(fill='Positive')) +
  labs(title='Dose',x="Antibody Levels [IgG]", y="Number of Samples", fill="Result") +
  scale_fill_manual(values = phs_colours(c("phs-blue", "phs-purple"))) +
  theme(aspect.ratio = 0.5)  +
  scale_y_log10() +
  scale_x_log10() +
  facet_wrap(. ~ stage)
```

```{r}
ggplot(data.bd , aes(x=IgG,fill=as.factor(QualResult))) +
  geom_histogram(position = "stack", bins=30) +
  #geom_histogram(position = "stack", bins=30, data=subset(data.bd, QualResult == 'Negative'), aes(fill='Negative')) +
  #geom_histogram(position = "stack", bins=30, data=subset(data.bd, QualResult == 'Equivocal'), aes(fill='Equivocal')) +
  geom_histogram(position = "stack", bins=30, data=subset(data.bd, QualResult == 'Positive'), aes(fill='Positive')) +
  labs(title='Dose',x="Antibody Levels [IgG]", y="Number of Samples", fill="Result") +
  scale_fill_manual(values = phs_colours(c("phs-magenta","phs-blue", "phs-purple"))) +
  theme(aspect.ratio = 0.5)  +
  scale_y_log10() +
  scale_x_log10() +
  facet_wrap(. ~ stage)
```






## Model A

Build analysis dataframes

```{r}
pc.modelA.df <- eavehelpers::get_modelA_df(data.pc)
bd.modelA.df <- eavehelpers::get_modelA_df(data.bd)
```




```{r,fig.height = 15, fig.width = 11}
pc.ages <- pc.modelA.df %>% select(age,Sex) %>% group_by(age,Sex) %>%
                 summarise(Pop=n()) %>% arrange(Pop,Sex)

ggplot(pc.ages, 
       mapping = aes(x = ifelse(Sex == "M", -Pop, Pop), 
                     y = as.factor(age), fill = Sex)) +
  scale_fill_discrete_phs(palette='all') + 
    scale_x_symmetric(labels = abs) +
  labs(x='Number of studies',y='Age') +
  geom_col() 

```


Perform a check on the confounding variables we want to use for modelA... 
This function returns the names of the variables that pass the univariable logit glm test fit for the outcome variable
```{r}
pc.modelA.vars <- pc.modelA.df %>% 
                  select(insufficient_response,prior_infection,
                         simd2020v2_sc_quintile,immuno,shielding,ch_resident) %>% 
                  eavehelpers::get_vars_to_use(outcome_var='insufficient_response')
print (pc.modelA.vars)
```


Doing this for the blood donors shows how it helps to remove the usage of the ch_resident variable (probably expected for carehome residents)
```{r}
bd.modelA.vars <- bd.modelA.df %>% 
                  select(insufficient_response,prior_infection,
                         simd2020v2_sc_quintile,immuno,shielding,ch_resident) %>% 
                  eavehelpers::get_vars_to_use(outcome_var='insufficient_response')
print (bd.modelA.vars)
```


```{r}
pc.modelA.gam.formula <- paste0(
  "insufficient_response ~ ", #outcome varialbe
  " s(age,by=cat) + cat", #age is a spline stratified by cat (product-dose categories)
  " + s(days_since_first_measurement) ", #spline of the number of days since the start of the pandemic/first-measurement
  " + s(days_since_last_vac) ", #days since the last vaccination, defined by the stage the measurement was taken at
  " + n_risk_gps ", #main exposure - number of QCOVID risk groups
  " + Sex + Q_BMI ", # confounding variables for Sex and BMI
  " + ",paste0(pc.modelA.vars,collapse = " + ") #other confounding from variables named and checked above
  ) 

pc.modelA.df <- pc.modelA.df %>% select(stage,all.vars(as.formula(pc.modelA.gam.formula)))
pc.modelA.gam.formula
```


```{r}
bd.modelA.gam.formula <- paste0(
  "insufficient_response ~ ", #outcome varialbe
  " s(age,by=cat) + cat", #age is a spline stratified by cat (product-dose categories)
  " + s(days_since_first_measurement) ", #spline of the number of days since the start of the pandemic/first-measurement
  " + s(days_since_last_vac) ", #days since the last vaccination, defined by the stage the measurement was taken at
  " + n_risk_gps ", #main exposure - number of QCOVID risk groups
  " + Sex + Q_BMI ", # confounding variables for Sex and BMI
  " + simd2020v2_sc_quintile",#force the inclusion of SIMD actually..
  " + ",paste0(bd.modelA.vars,collapse = " + ") #other confounding from variables named and checked above
  ) 

bd.modelA.df <- bd.modelA.df %>% select(stage,all.vars(as.formula(bd.modelA.gam.formula)))
bd.modelA.gam.formula
```



```{r}
pc.modelA.gamFit <- eavehelpers::perform_gam(pc.modelA.df,as.formula(pc.modelA.gam.formula))
summary(pc.modelA.gamFit)
```


```{r}
bd.modelA.gamFit <- eavehelpers::perform_gam(bd.modelA.df,as.formula(bd.modelA.gam.formula))
summary(bd.modelA.gamFit)
```



```{r}
pc.modelA.gamFit.or <- eavehelpers::get_or_from_gam(pc.modelA.gamFit) 
pc.modelA.gamFit.or %>% select(label,level,OR,LCL,UCL)
```



```{r}
bd.modelA.gamFit.or <- eavehelpers::get_or_from_gam(bd.modelA.gamFit) %>% filter(is.finite(UCL))
bd.modelA.gamFit.or %>% select(label,level,OR,LCL,UCL)
```


```{r,fig.height = 9, fig.width = 9}
eavehelpers::plot_ratios(pc.modelA.gamFit.or %>% rename(Nominal=OR))
```


```{r,fig.height = 9, fig.width = 9}
eavehelpers::plot_ratios(bd.modelA.gamFit.or %>% rename(Nominal=OR))
```

```{r}
modelA.gamFit.or <- pc.modelA.gamFit.or %>% mutate(cohort='Primary Care') %>% rbind(bd.modelA.gamFit.or %>% mutate(cohort='Blood Donors') ) %>%
                    rename(Nominal=OR)
saveRDS(modelA.gamFit.or,paste0(output_folder,'/modelA_gamFit_OR.rds'))

eavehelpers::display_ratios_table(modelA.gamFit.or)
```



```{r,fig.height = 9, fig.width = 11}

p.modelA.gamFit.or <- eavehelpers::plot_ratios(modelA.gamFit.or,
                                               fill=cohort,
                                               facet=as.factor(label) ~ fct_rev(as.factor(cohort))
                                               ) 
p.modelA.gamFit.or
```



## Model B

Build analysis dataframes

```{r}
pc.modelB.df <- eavehelpers::get_modelB_df(data.pc)
bd.modelB.df <- eavehelpers::get_modelB_df(data.bd)
```

Check these variables again...
```{r}
pc.modelB.vars <- pc.modelB.df %>% 
                  select(insufficient_response,prior_infection,
                         simd2020v2_sc_quintile,immuno,shielding,ch_resident) %>% 
                  eavehelpers::get_vars_to_use(outcome_var='insufficient_response')
print (pc.modelB.vars)
```

```{r}
bd.modelB.vars <- bd.modelB.df %>% 
                  select(insufficient_response,prior_infection,
                         simd2020v2_sc_quintile,immuno,shielding,ch_resident) %>% 
                  eavehelpers::get_vars_to_use(outcome_var='insufficient_response')
print (bd.modelB.vars)
```


Use the helper function to work out which QCOVID terms to include. 
We use ones where there are at least 5 events with the outcome of insufficient response
```{r}
pc.modelB.qnames <- get_qcovid_names(pc.modelB.df,n=5)
pc.modelB.qnames
```

```{r}
bd.modelB.qnames <- get_qcovid_names(bd.modelB.df,n=5)
bd.modelB.qnames
```

```{r}
pc.modelB.gam.formula <- paste0(
  "insufficient_response ~ ", #outcome varialbe
  " s(age,by=cat) + cat", #age is a spline stratified by cat (product-dose categories)
  " + s(days_since_first_measurement) ", #spline of the number of days since the start of the pandemic/first-measurement
  " + s(days_since_last_vac) ", #days since the last vaccination, defined by the stage the measurement was taken at
  " + ",paste0(pc.modelB.qnames,collapse = " + "), #main exposure - the individiual qcovid risk groups
  " + Sex + Q_BMI ", # confounding variables for Sex and BMI
  " + ",paste0(pc.modelB.vars,collapse = " + ") #other confounding from variables named and checked above
  ) 

#pc.modelB.df <- pc.modelB.df %>% select(stage,all.vars(as.formula(pc.modelB.gam.formula)))
pc.modelB.gam.formula
```


```{r}
bd.modelB.gam.formula <- paste0(
  "insufficient_response ~ ", #outcome varialbe
  " s(age,by=cat) + cat", #age is a spline stratified by cat (product-dose categories)
  " + s(days_since_first_measurement) ", #spline of the number of days since the start of the pandemic/first-measurement
  " + s(days_since_last_vac) ", #days since the last vaccination, defined by the stage the measurement was taken at
  " + ",paste0(bd.modelB.qnames,collapse = " + "), #main exposure - the individiual qcovid risk groups
  " + Sex + Q_BMI ", # confounding variables for Sex and BMI
  " + simd2020v2_sc_quintile",#force the inclusion of SIMD actually..
  " + ",paste0(bd.modelB.vars,collapse = " + ") #other confounding from variables named and checked above
  ) 

#bd.modelA.df <- bd.modelA.df %>% select(stage,all.vars(as.formula(bd.modelA.gam.formula)))
bd.modelB.gam.formula
```



```{r}
pc.modelB.gamFit <- eavehelpers::perform_gam(pc.modelB.df,as.formula(pc.modelB.gam.formula))
summary(pc.modelB.gamFit)
```


```{r}
bd.modelB.gamFit <- eavehelpers::perform_gam(bd.modelB.df,as.formula(bd.modelB.gam.formula))
summary(bd.modelB.gamFit)
```



```{r}
pc.modelB.gamFit.or <- eavehelpers::get_or_from_gam(pc.modelB.gamFit) 
pc.modelB.gamFit.or %>% select(label,level,OR,LCL,UCL)
```


```{r,fig.height = 15, fig.width = 9}
eavehelpers::plot_ratios(pc.modelB.gamFit.or %>% rename(Nominal=OR))
```


```{r,fig.height = 12, fig.width = 9}
eavehelpers::plot_ratios(pc.modelB.gamFit.or %>% rename(Nominal=OR) %>% 
                         mutate(
                           level=ifelse(grepl("Q_DIAG",names) & level=='Yes',label,level),
                           label=ifelse(grepl("Q_DIAG",names),'QCOVID',label)
                         )
)
                         
```



```{r}
bd.modelB.gamFit.or <- eavehelpers::get_or_from_gam(bd.modelB.gamFit) %>% filter(is.finite(UCL))
bd.modelB.gamFit.or %>% filter(var=='cat') %>% select(label,level,OR,LCL,UCL) 
```


```{r}
bd.modelB.df %>% filter(stage<3) %>% group_by(cat,insufficient_response) %>% summarise(n=n())
```

```{r}
temp %>% filter(stage<3 & n_risk_gps>0) %>% group_by(cat,outcome) %>% summarise(n=n())
```

```{r}
bd.modelB.df %>% ggplot(aes(x=age)) + geom_histogram()
```


```{r}
bd.modelB.gam.formula
```


```{r,fig.height = 3, fig.width = 9}

f <- "insufficient_response ~ s(age) +  cat "

bd.modelB.gamFit <- eavehelpers::perform_gam(bd.modelB.df
                                             ,as.formula(f))
bd.modelB.gamFit.or <- eavehelpers::get_or_from_gam(bd.modelB.gamFit) 

eavehelpers::plot_ratios(bd.modelB.gamFit.or %>% rename(Nominal=OR))
```

```{r}
plot_gam(bd.modelB.gamFit,'age',var_ref=40,list(cat=levels((bd.modelB.df%>% filter(stage<3))$cat))) + 
  facet_grid(~cat,scales="free") +  
  coord_flip() + scale_y_log10(labels = ~ signif(.x,digits=1)) + # limits=c(0.11,11) ) + 
  theme(strip.background.x = element_blank(),
        strip.text.x = element_blank(),
        legend.position="top")

```





```{r}
modelB.gamFit.or <- pc.modelB.gamFit.or %>% mutate(cohort='Primary Care') %>% rbind(bd.modelB.gamFit.or %>% mutate(cohort='Blood Donors') ) %>%
                    rename(Nominal=OR)
saveRDS(modelB.gamFit.or,paste0(output_folder,'/modelB_gamFit_OR.rds'))

eavehelpers::display_ratios_table(modelB.gamFit.or)
```





```{r,fig.height = 9, fig.width = 13}


p.modelB.gamFit.or <- eavehelpers::plot_ratios(modelB.gamFit.or %>% 
                                                  mutate(
                                                         level=ifelse(grepl("Q_DIAG",names) & level=='Yes',label,level),
                                                         label=ifelse(grepl("Q_DIAG",names),'QCOVID',label)
                                                       ),
                                               fill=cohort,
                                               facet=as.factor(label) ~ fct_rev(as.factor(cohort))
                                               ) 
p.modelB.gamFit.or
```














