---
title: "Insufficient Response Analysis"
output: html_notebook
---



## Initial setup

```{r}
library(eavehelpers)
library(ggplot2)
library(phsstyles)



```


### Load Data

See notebook called 'LoadData.Rmd' for how the analysis dataframes were created:

```{r}
data.pc <- readRDS("/conf/EAVE/GPanalysis/data/temp/fail_to_mount_pc_3.rds")
data.bd <- readRDS("/conf/EAVE/GPanalysis/data/temp/fail_to_mount_bd_3.rds")
print (nrow(data.pc))
print (nrow(data.bd))
```


Create an output folder for the date we are running the analysis.

```{r}
output_folder <- file.path("results", Sys.Date())
dir.create(output_folder, showWarnings = TRUE)
```


## Model A

Build analysis dataframes

```{r}
pc.modelA.df <- eavehelpers::get_modelA_df(data.pc)
bd.modelA.df <- eavehelpers::get_modelA_df(data.bd)
```


### Get variables to use in the models


Perform a check on the confounding variables we want to use for modelA... 
This function returns the names of the variables that pass the univariable logit glm test fit for the outcome variable
```{r}
pc.modelA.vars <- pc.modelA.df %>% 
                  select(insufficient_response,prior_infection,
                         simd2020v2_sc_quintile,immuno,shielding,ch_resident) %>% 
                  eavehelpers::get_vars_to_use(outcome_var='insufficient_response')
print (pc.modelA.vars)
```


Doing this for the blood donors shows how it helps to remove the usage of the ch_resident variable (probably expected for carehome residents)
```{r}
bd.modelA.vars <- bd.modelA.df %>% 
                  select(insufficient_response,prior_infection,
                         simd2020v2_sc_quintile,immuno,shielding,ch_resident) %>% 
                  eavehelpers::get_vars_to_use(outcome_var='insufficient_response')
print (bd.modelA.vars)
```

### Construct formulas for the models


```{r}
pc.modelA.gam.formula <- paste0(
  "insufficient_response ~ ", #outcome varialbe
  " s(age,by=cat,k=5) + cat", #age is a spline stratified by cat (product-dose categories)
  " + s(days_since_first_measurement) ", #spline of the number of days since the start of the pandemic/first-measurement
  " + s(days_since_last_vac) ", #days since the last vaccination, defined by the stage the measurement was taken at
  " + n_risk_gps ", #main exposure - number of QCOVID risk groups
  " + Sex + Q_BMI ", # confounding variables for Sex and BMI
  " + ",paste0(pc.modelA.vars,collapse = " + ") #other confounding from variables named and checked above
  ) 

pc.modelA.df <- pc.modelA.df %>% select(stage,all.vars(as.formula(pc.modelA.gam.formula)))
pc.modelA.gam.formula
```


```{r}
bd.modelA.gam.formula <- paste0(
  "insufficient_response ~ ", #outcome varialbe
  " s(age,by=cat,k=5) + cat", #age is a spline stratified by cat (product-dose categories)
  " + s(days_since_first_measurement) ", #spline of the number of days since the start of the pandemic/first-measurement
  " + s(days_since_last_vac) ", #days since the last vaccination, defined by the stage the measurement was taken at
  " + n_risk_gps ", #main exposure - number of QCOVID risk groups
  " + Sex + Q_BMI ", # confounding variables for Sex and BMI
  " + simd2020v2_sc_quintile",#force the inclusion of SIMD actually..
  " + ",paste0(bd.modelA.vars,collapse = " + ") #other confounding from variables named and checked above
  ) 

bd.modelA.df <- bd.modelA.df %>% select(stage,all.vars(as.formula(bd.modelA.gam.formula)))
bd.modelA.gam.formula
```


### Fit the GAMs

```{r}
pc.modelA.gamFit <- eavehelpers::perform_gam(pc.modelA.df,as.formula(pc.modelA.gam.formula))
summary(pc.modelA.gamFit)
```

Use a helper function to perform the GAM fit unadjusted (each variable individually)
```{r}
attr(pc.modelA.gamFit,'unadjusted') <- eavehelpers::calculate_unadjusted_gam(pc.modelA.df,
                                                                              as.formula(pc.modelA.gam.formula))
labels(attributes(pc.modelA.gamFit)$unadjusted)
```

No the adjusted odds ratios for model A..
```{r}
bd.modelA.gamFit <- eavehelpers::perform_gam(bd.modelA.df,as.formula(bd.modelA.gam.formula))
summary(bd.modelA.gamFit)
```

Get the unadjusted numbers for the blood donors
```{r}
attr(bd.modelA.gamFit,'unadjusted') <- eavehelpers::calculate_unadjusted_gam(bd.modelA.df,
                                                                              as.formula(bd.modelA.gam.formula))
labels(attributes(bd.modelA.gamFit)$unadjusted)
```



### Save the fitted models 

Make a temporary save of the GAM objects incase we ever need to open them for future comparisons/checks.
```{r}
saveRDS(pc.modelA.gamFit,paste0(output_folder,'/modelA_pc_gamFit.rds'))
saveRDS(bd.modelA.gamFit,paste0(output_folder,'/modelA_bd_gamFit.rds'))
```


### ORs (linear variables)

Extract the Odds Ratios (ORs) for the parametric variables (linear variables) used from the GAM
```{r}
pc.modelA.gamFit.or <- eavehelpers::get_or_from_gam(pc.modelA.gamFit) 
pc.modelA.gamFit.or %>% select(label,level,OR,uOR)
```



```{r}
bd.modelA.gamFit.or <- eavehelpers::get_or_from_gam(bd.modelA.gamFit) %>% filter(is.finite(UCL))
bd.modelA.gamFit.or %>% select(level,OR,uOR)
```


```{r,fig.height = 9, fig.width = 9}
eavehelpers::plot_ratios(pc.modelA.gamFit.or %>% rename(Nominal=OR))
```


```{r,fig.height = 9, fig.width = 9}
eavehelpers::plot_ratios(bd.modelA.gamFit.or %>% rename(Nominal=OR))
```

```{r}
modelA.gamFit.or <- pc.modelA.gamFit.or %>% mutate(cohort='Primary Care') %>% rbind(bd.modelA.gamFit.or %>% mutate(cohort='Blood Donors') ) %>% rename(Nominal=OR)
saveRDS(modelA.gamFit.or,paste0(output_folder,'/modelA_gamFit_OR.rds'))

eavehelpers::display_ratios_table(modelA.gamFit.or)
```



```{r,fig.height = 9, fig.width = 11}

p.modelA.gamFit.or <- eavehelpers::plot_ratios(modelA.gamFit.or,
                                               fill=cohort,
                                               facet=as.factor(label) ~ fct_rev(as.factor(cohort))
                                               ) 
p.modelA.gamFit.or
```



## Model B

Build analysis dataframes

```{r}
pc.modelB.df <- eavehelpers::get_modelB_df(data.pc)
bd.modelB.df <- eavehelpers::get_modelB_df(data.bd)
```

Check these variables again...
```{r}
pc.modelB.vars <- pc.modelB.df %>% 
                  select(insufficient_response,prior_infection,
                         simd2020v2_sc_quintile,immuno,shielding,ch_resident) %>% 
                  eavehelpers::get_vars_to_use(outcome_var='insufficient_response')
print (pc.modelB.vars)
```

```{r}
bd.modelB.vars <- bd.modelB.df %>% 
                  select(insufficient_response,prior_infection,
                         simd2020v2_sc_quintile,immuno,shielding,ch_resident) %>% 
                  eavehelpers::get_vars_to_use(outcome_var='insufficient_response')
print (bd.modelB.vars)
```


Use the helper function to work out which QCOVID terms to include. 
We use ones where there are at least 5 events with the outcome of insufficient response
```{r}
pc.modelB.qnames <- get_qcovid_names(pc.modelB.df,n=5)
pc.modelB.qnames
```

```{r}
bd.modelB.qnames <- get_qcovid_names(bd.modelB.df,n=5)
bd.modelB.qnames
```

```{r}
pc.modelB.gam.formula <- paste0(
  "insufficient_response ~ ", #outcome varialbe
  " s(age,by=cat) + cat", #age is a spline stratified by cat (product-dose categories)
  " + s(days_since_first_measurement) ", #spline of the number of days since the start of the pandemic/first-measurement
  " + s(days_since_last_vac) ", #days since the last vaccination, defined by the stage the measurement was taken at
  " + ",paste0(pc.modelB.qnames,collapse = " + "), #main exposure - the individiual qcovid risk groups
  " + Sex + Q_BMI ", # confounding variables for Sex and BMI
  " + simd2020v2_sc_quintile",#force the inclusion of SIMD actually..
  " + ",paste0(pc.modelB.vars,collapse = " + ") #other confounding from variables named and checked above
  ) 

#pc.modelB.df <- pc.modelB.df %>% select(stage,all.vars(as.formula(pc.modelB.gam.formula)))
pc.modelB.gam.formula
```


```{r}
bd.modelB.gam.formula <- paste0(
  "insufficient_response ~ ", #outcome varialbe
  " s(age,by=cat,k=3) + cat", #age is a spline stratified by cat (product-dose categories)
  " + s(days_since_first_measurement) ", #spline of the number of days since the start of the pandemic/first-measurement
  " + s(days_since_last_vac) ", #days since the last vaccination, defined by the stage the measurement was taken at
  " + ",paste0(bd.modelB.qnames,collapse = " + "), #main exposure - the individiual qcovid risk groups
  " + Sex + Q_BMI ", # confounding variables for Sex and BMI
  " + simd2020v2_sc_quintile",#force the inclusion of SIMD actually..
  " + ",paste0(bd.modelB.vars,collapse = " + ") #other confounding from variables named and checked above
  ) 

#bd.modelA.df <- bd.modelA.df %>% select(stage,all.vars(as.formula(bd.modelA.gam.formula)))
bd.modelB.gam.formula
```



```{r}
pc.modelB.gamFit <- eavehelpers::perform_gam(pc.modelB.df,as.formula(pc.modelB.gam.formula))
summary(pc.modelB.gamFit)
```


```{r}
attr(pc.modelB.gamFit,'unadjusted') <- eavehelpers::calculate_unadjusted_gam(pc.modelB.df,
                                                                              as.formula(pc.modelB.gam.formula))
labels(attributes(pc.modelB.gamFit)$unadjusted)
```

```{r}
bd.modelB.gamFit <- eavehelpers::perform_gam(bd.modelB.df,as.formula(bd.modelB.gam.formula))
summary(bd.modelB.gamFit)
```


```{r}
attr(bd.modelB.gamFit,'unadjusted') <- eavehelpers::calculate_unadjusted_gam(bd.modelB.df,
                                                                              as.formula(bd.modelB.gam.formula))
labels(attributes(bd.modelB.gamFit)$unadjusted)
```



### Save the fitted models 

Make a temporary save of the GAM objects incase we ever need to open them for future comparisons/checks.
```{r}
saveRDS(pc.modelB.gamFit,paste0(output_folder,'/modelB_pc_gamFit.rds'))
saveRDS(bd.modelB.gamFit,paste0(output_folder,'/modelB_bd_gamFit.rds'))
```




```{r}
pc.modelB.gamFit.or <- eavehelpers::get_or_from_gam(pc.modelB.gamFit) 
pc.modelB.gamFit.or %>% select(label,level,OR,LCL,UCL)
```


```{r,fig.height = 15, fig.width = 9}
eavehelpers::plot_ratios(pc.modelB.gamFit.or %>% rename(Nominal=OR))
```


```{r,fig.height = 12, fig.width = 9}
eavehelpers::plot_ratios(pc.modelB.gamFit.or %>% rename(Nominal=OR) %>% 
                         mutate(
                           level=ifelse(grepl("Q_DIAG",names) & level=='Yes',label,level),
                           label=ifelse(grepl("Q_DIAG",names),'QCOVID',label)
                         )
)
                         
```



```{r}
bd.modelB.gamFit.or <- eavehelpers::get_or_from_gam(bd.modelB.gamFit) %>% filter(is.finite(UCL))
bd.modelB.gamFit.or %>% select(label,level,OR,LCL,UCL) 
```



```{r}
#plot_gam(bd.modelB.gamFit,'age',var_ref=40,list(cat=levels((bd.modelB.df%>% filter(stage<3))$cat))) + 
#  facet_grid(~cat,scales="free") +  
#  coord_flip() + scale_y_log10(labels = ~ signif(.x,digits=1)) + # limits=c(0.11,11) ) + 
#  theme(strip.background.x = element_blank(),
#        strip.text.x = element_blank(),
#        legend.position="top")

```





```{r}
modelB.gamFit.or <- pc.modelB.gamFit.or %>% mutate(cohort='Primary Care') %>% 
                    rbind(bd.modelB.gamFit.or %>% 
                         mutate(cohort='Blood Donors') ) %>%
                    rename(Nominal=OR)
saveRDS(modelB.gamFit.or,paste0(output_folder,'/modelB_gamFit_OR.rds'))

eavehelpers::display_ratios_table(modelB.gamFit.or)
```





```{r,fig.height = 9, fig.width = 13}


p.modelB.gamFit.or <- eavehelpers::plot_ratios(modelB.gamFit.or %>% 
                                                  mutate(
                                                         level=ifelse(grepl("Q_DIAG",names) & level=='Yes',label,level),
                                                         label=ifelse(grepl("Q_DIAG",names),'QCOVID',label)
                                                       ),
                                               fill=cohort,
                                               facet=as.factor(label) ~ fct_rev(as.factor(cohort))
                                               ) 
p.modelB.gamFit.or
```














