---
title: "Insufficient Response Analysis"
output: html_notebook
---


### Setup

Import some libraries we need.

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(phsstyles)
library(lemon)
library(eavehelpers)
```

### Load Data

Use the helper function to load up the recommended, stable datasets needed for this analysis
```{r}
eave.data <- eavehelpers::load_recommended_datasets() 
labels(eave.data)
```

create two dataframes that join the two serology datasets up with other EAVE-II files
```{r}
data.pc <- eavehelpers::serology_vaccine_analysis.create_dataframe(eave.data,'serology_primary_care')
data.bd <- eavehelpers::serology_vaccine_analysis.create_dataframe(eave.data,'serology_blood_donors')
```

```{r}
attributes(data.pc)$meta
```


```{r}
attributes(data.bd)$meta
```

### Save Data

Save the analysis dataframe so we can:
1. Bookkeep the versions of the data we are using, incase changes are detected, or input files are corrupted.
2. Not have to construct the analysis dataframe everytime we run the code
3. Be able to easily re-run the results once the input datasets are fixed

```{r}
saveRDS(data.pc,"/conf/EAVE/GPanalysis/data/temp/fail_to_mount_pc_Jan10_v4.rds")
saveRDS(data.bd,"/conf/EAVE/GPanalysis/data/temp/fail_to_mount_bd_Jan10_v4.rds")
```


### Simple data checks


#### Primary care data

```{r}
ggplot(data.pc , aes(x=IgG,fill=as.factor(QualResult))) +
  #geom_histogram(position = "stack", bins=30) +
  geom_histogram(position = "stack", bins=30, data=subset(data.pc, QualResult == 'Negative'), aes(fill='Negative')) +
  #geom_histogram(position = "stack", bins=30, data=subset(data.pc, QualResult == 'Equivocal'), aes(fill='Equivocal')) +
  geom_histogram(position = "stack", bins=30, data=subset(data.pc, QualResult == 'Positive'), aes(fill='Positive')) +
  labs(title='Dose',x="Antibody Levels [IgG]", y="Number of Samples", fill="Result") +
  scale_fill_manual(values = phs_colours(c("phs-blue", "phs-purple"))) +
  theme(aspect.ratio = 0.5)  +
  scale_y_log10() +
  scale_x_log10() +
  facet_wrap(. ~ stage)
```

#### Blood donors data


```{r}
ggplot(data.bd , aes(x=IgG,fill=as.factor(QualResult))) +
  geom_histogram(position = "stack", bins=30) +
  #geom_histogram(position = "stack", bins=30, data=subset(data.bd, QualResult == 'Negative'), aes(fill='Negative')) +
  #geom_histogram(position = "stack", bins=30, data=subset(data.bd, QualResult == 'Equivocal'), aes(fill='Equivocal')) +
  geom_histogram(position = "stack", bins=30, data=subset(data.bd, QualResult == 'Positive'), aes(fill='Positive')) +
  labs(title='Dose',x="Antibody Levels [IgG]", y="Number of Samples", fill="Result") +
  scale_fill_manual(values = phs_colours(c("phs-magenta","phs-blue", "phs-purple"))) +
  theme(aspect.ratio = 0.5)  +
  scale_y_log10() +
  scale_x_log10() +
  facet_wrap(. ~ stage)
```




```{r,fig.height = 5, fig.width = 8}
pc.ages <- data.pc %>% mutate(age = cut(ageYear,seq(0,100,10))) %>% filter(!is.na(age)) %>% 
                   select(age,Sex) %>% group_by(age,Sex) %>%
                   summarise(Pop=n()) %>% arrange(Pop,Sex)

ggplot(pc.ages, 
       mapping = aes(x = ifelse(Sex == "M", -Pop, Pop), 
                     y = as.factor(age), fill = Sex)) +
  scale_fill_discrete_phs(palette='all') + 
    scale_x_symmetric(labels = abs) +
  labs(x='Number of studies',y='Age') +
  geom_col() 

```

```{r,fig.height = 5, fig.width = 8}
bd.ages <- data.bd %>% mutate(age = cut(ageYear,seq(0,100,10))) %>% filter(!is.na(age)) %>% 
                   select(age,Sex) %>% group_by(age,Sex) %>%
                   summarise(Pop=n()) %>% arrange(Pop,Sex)

ggplot(bd.ages, 
       mapping = aes(x = ifelse(Sex == "M", -Pop, Pop), 
                     y = as.factor(age), fill = Sex)) +
  scale_fill_discrete_phs(palette='all') + 
    scale_x_symmetric(labels = abs) +
  labs(x='Number of studies',y='Age') +
  geom_col() 

```

