---
title: "Severe Outcomes Analysis"
output: html_notebook
---



## Initial setup

```{r}
library(eavehelpers)
library(ggplot2)
library(phsstyles)
library(survival)

data.pc <- readRDS("/conf/EAVE/GPanalysis/data/temp/fail_to_mount_pc_Jan13_final.rds")
data.bd <- readRDS("/conf/EAVE/GPanalysis/data/temp/fail_to_mount_bd_Jan13_final.rds")
print (nrow(data.pc))
print (nrow(data.bd))
```

Create an output folder for the date we are running the analysis.

```{r}
output_folder <- file.path("results", "FINAL")# Sys.Date())
dir.create(output_folder, showWarnings = TRUE)
```



```{r}
#eave.data <- eavehelpers::load_recommended_datasets() 
labels(eave.data)
```


```{r}
pc.modelC.df <- eavehelpers::serology_outcome_analysis.create_infect_dataframe(data.pc,eave.data) %>% code_vars()
bd.modelC.df <- eavehelpers::serology_outcome_analysis.create_infect_dataframe(data.bd,eave.data) %>% code_vars()
```


```{r}
pc.modelC.cox <- with(pc.modelC.df,
                 coxph(Surv(t1,t2, outcome_infect) ~  insufficient_response + n_risk_gps + ageYear + Sex + shielding + ch_resident + Q_BMI + simd2020v2_sc_quintile + ur6_2016_name + additional ))

pc.modelC.cox.HRs <- eavehelpers::get_forest_df(pc.modelC.cox,pc.modelC.df)

eavehelpers::display_ratios_table(pc.modelC.cox.HRs %>% mutate(Nominal=estimate,LCL=conf.low,UCL=conf.high,label=name))
```


```{r}
pc.modelC.cox.unadjusted <- with(pc.modelC.df,
                 coxph(Surv(t1,t2, outcome_infect) ~  insufficient_response ))

pc.modelC.cox.uHRs <- eavehelpers::get_forest_df(pc.modelC.cox.unadjusted,pc.modelC.df)

attr(pc.modelC.cox.HRs,'unadjusted') <- pc.modelC.cox.uHRs

eavehelpers::display_ratios_table( pc.modelC.cox.uHRs %>% mutate(Nominal=estimate,LCL=conf.low,UCL=conf.high,label=name))
```



```{r}
bd.modelC.cox <- with(bd.modelC.df,
                 coxph(Surv(t1,t2, outcome_infect) ~  insufficient_response + n_risk_gps + ageYear + Sex + shielding  + Q_BMI + simd2020v2_sc_quintile + ur6_2016_name + additional ))
summary(bd.modelC.cox)
```


```{r}
bd.modelC.cox.HRs <- eavehelpers::get_forest_df(bd.modelC.cox,bd.modelC.df)
eavehelpers::display_ratios_table(bd.modelC.cox.HRs %>% mutate(Nominal=estimate,LCL=conf.low,UCL=conf.high,label=name))
```


```{r}
bd.modelC.cox.unadjusted <- with(bd.modelC.df,
                 coxph(Surv(t1,t2, outcome_infect) ~  insufficient_response ))

bd.modelC.cox.uHRs <- eavehelpers::get_forest_df(bd.modelC.cox.unadjusted,bd.modelC.df)

attr(bd.modelC.cox.HRs,'unadjusted') <- bd.modelC.cox.uHRs

eavehelpers::display_ratios_table( bd.modelC.cox.uHRs %>% mutate(Nominal=estimate,LCL=conf.low,UCL=conf.high,label=name))
```



```{r,fig.height = 9, fig.width = 9}

modelC.cox.HRs <- pc.modelC.cox.HRs %>% mutate(dataset='Primary Care') %>% 
                           rbind(bd.modelC.cox.HRs %>% mutate(dataset='Blood Donors')) %>%
                 mutate(dataset=as.factor(dataset)) %>%
                 mutate(Nominal=estimate,UCL=conf.high,LCL=conf.low)

eavehelpers::plot_ratios(modelC.cox.HRs,
                         xmin=0.25,xmax=5,
                         xtitle='HR (95% CI)',
                         fill=dataset,
                         facet=as.factor(name) ~ fct_rev(as.factor(dataset))
) 
```


## Model D


Create the analysis ready dataframe that can be used for both model D and model E.
* The filter of `ageYear!='0-19'` is because there is one event in this category, which messes up the fit. We require >5 events in all categories. This is just a simple  ad-hoc way around this, instead of regrouping and/or redefining categories.

```{r}
pc.modelDE.df <- eavehelpers::serology_outcome_analysis.create_hosp_death_dataframe(data.pc,eave.data) %>% code_vars() %>% filter(ageYear!='0-19') %>% droplevels()
bd.modelDE.df <- eavehelpers::serology_outcome_analysis.create_hosp_death_dataframe(data.bd,eave.data) %>% code_vars() %>% filter(ageYear!='0-19') %>% droplevels()
```


```{r}
pc.modelD.cox <- with(pc.modelDE.df,
                          coxph(Surv(t1,t2, outcome_hosp) ~ insufficient_response + additional + n_risk_gps + ageYear + Sex + Q_BMI  ))
summary(pc.modelD.cox)
```



```{r}
pc.modelD.cox.HRs <- eavehelpers::get_forest_df(pc.modelD.cox,pc.modelDE.df)
eavehelpers::display_ratios_table( pc.modelD.cox.HRs %>% mutate(Nominal=estimate,LCL=conf.low,UCL=conf.high,label=name))
```

```{r}
eavehelpers::plot_ratios(pc.modelD.cox.HRs %>% mutate(Nominal=estimate,LCL=conf.low,UCL=conf.high,label=name),
                         xtitle='HR (95% CI)')
```


```{r}
f.pc <- base::c('Undetectable'=0.,'Insufficient'=4.82,'Below Average'=33.8,'Average'=230.,'Above Average'=2000)
f.bd <- base::c('Undetectable'=-1.,'Insufficient'=0.11,'Below Average'=0.7,'Average'=5,'Above Average'=8.5)

pc.modelE.df <- pc.modelDE.df 

f <- f.pc
labels <- labels(f)
calculate_igg_quartiles <- function(x){
  return (labels(f[sum(f<=x)])[1])
}

pc.modelE.df$igg <- mapply(calculate_igg_quartiles, pc.modelE.df$IgG)
pc.modelE.df <- pc.modelE.df %>% mutate(igg = factor(igg,levels=labels)) %>% mutate(igg=fct_relevel(igg,'Average'))
pc.modelE.df %>% group_by(igg,outcome_hosp) %>% summarise(n=n())

```

Calculate a formular that will calculate the number of risks groups expect for four specific risk groups
```{r}
qnames <- pc.modelE.df %>% eavehelpers::get_qcovid_names(var='outcome_hosp',condition=1,n=1)
#qnames2 <- c("Q_DIAG_BLOOD_CANCER","Q_DIAG_CHD","Q_DIAG_PULM_HYPER","Q_DIAG_COPD")
qnames2 <- c("Q_DIAG_BLOOD_CANCER","Q_DIAG_CHD","Q_DIAG_PULM_HYPER","Q_DIAG_CKD")

formula_other <- paste0(paste0(qnames,collapse = ' + ')," - ",paste0(qnames2,collapse = ' - '))
formula_other
```


```{r}
pc.modelE.df <- pc.modelE.df %>%
  mutate_at(vars(contains("Q_DIAG")),~as.numeric(ifelse(.=='Yes',1,0))) %>%
  mutate(n_risk_gpsOther = !!parse_quo(formula_other, env = caller_env())) %>%
  mutate(n_risk_gpsOther = as.factor(ifelse(n_risk_gpsOther>2,'3+',n_risk_gpsOther))) %>% 
  mutate_at(vars(contains("Q_DIAG")),~as.factor(ifelse(.==1,'Yes','No'))) 


formula <- paste0("Surv(t1,t2, outcome_hosp) ~ igg + additional  + ",
                              paste0(qnames2,collapse=" + "),
                              "+ n_risk_gpsOther + ageYear + Sex + Q_BMI " )

pc.modelE.cox <- with(pc.modelE.df,coxph(as.formula(formula)))
summary(pc.modelE.cox)
```

```{r}
pc.modelE.cox.HRs <- eavehelpers::get_forest_df(pc.modelE.cox,pc.modelE.df)
eavehelpers::display_ratios_table( pc.modelE.cox.HRs %>% mutate(Nominal=estimate,LCL=conf.low,UCL=conf.high,label=name))
```

```{r,fig.height = 8, fig.width = 7}
eavehelpers::plot_ratios(pc.modelE.cox.HRs %>% mutate(label=name,Nominal=estimate,LCL=conf.low,UCL=conf.high),
                         xmin=0.1,xmax=10,
                         xtitle='HR (95% CI)'
)
```

```{r}
saveRDS(pc.modelC.cox.HRs,paste0(output_folder,'/modelC_pc_cox_HR.rds'))
saveRDS(bd.modelC.cox.HRs,paste0(output_folder,'/modelC_bd_cox_HR.rds'))
saveRDS(pc.modelD.cox.HRs,paste0(output_folder,'/modelD_cox_HR.rds'))
saveRDS(pc.modelE.cox.HRs,paste0(output_folder,'/modelE_cox_HR.rds'))
```


